---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# AgeEstimatoR
**Estimate age from otolith pictures using Deep Learning. R package version `r packageVersion("AgeEstimatoR")`**
<!-- badges: start -->
<!-- badges: end -->

This is an early draft of an R package to standardize fish otolith photographs and use 
Deep Learning (DL or ML) models to estimate the age from the standardized photographs. 
The package currently works only for Greenland halibut and uses the models 
developed by [Martinsen et al. (2022)](https://doi.org/10.1371/journal.pone.0277244) but 
may be extended for other species and models in the future. The age estimation
uses Python and the TensorFlow platform for machine learning. Note that the DL age
estimation method was developed only for **right otoliths of Greenland halibut**.  

Note that running the age estimation function (`estimate_age`) downloads Python and TensorFlow into a virtual environment together with DL models for Greenland halibut. These **take >2.2 gigabytes of file space** in total.

## Installation

You can install AgeEstimatoR using the remotes or devtools packages:

``` {r, eval = FALSE}
remotes::install_github("DeepWaterIMR/AgeEstimatoR")
```

## Example

```{r}
library(AgeEstimatoR)
```

### Image standardization

Standardize images ready for age estimation:

```{r, eval = FALSE}
standardize_images(
  input_path = file.path(system.file("extdata", package = "AgeEstimatoR"), 
                         "example_images", "input"),
  output_path = "inst/extdata/example_images/standardized/",
  adj = 1.1, fillparam = 3
)
```

The function brings up a plot screen where you'll need to mark the corners of the **right** otolith for cropping. To do the marking, click the photograph twice until you see a line:

Next, the resulting cropped photo will be displayed (left without filtering, right with filtering). You will no be asked whether you are satisfied with the result on the right. If you answer 2 (no), the function will ask new adjustment factor and fill parameter () and you will get a new shot in cropping the photo. If you answer 1 (yes), the function will jump to the next photo. 

### Age estimation

It is important to understand how the `estimate_age` function **can clutter your computer full of Python installations if used wrong**: The age estimation requires Python 3.10, TensorFlow and very specific versions of Python modules. Consequently, the `estimate_age()` function downloads complete Python 3.10, TensorFlow and required modules into a virtual environment folder called `python_virtualenv` by default in the project root. The folder takes 1.43 GB of file space and clutter your computer if using AgeEstimatoR in multiple projects. You can use the `venv_path` to define the path to the virtual environment once you have downloaded it once. 

Use the `data_path` argument to save the results directly into a file. 

```{r, eval = FALSE}
x <- estimate_age(
  image_path = file.path(system.file("extdata", package = "AgeEstimatoR"), 
                         "example_images", "standardized")
  )
```


## Things that remain to be solved

- Is the machine learning script (`inst/python/dl_age_estimator.py`) correct?
  - Is the order of sexes in the output correct?
- File size of standardized photographs is only 10-20 KB. Is this a problem?
- Automatic otolith recognition to skip that otolith marking phase
- More plotting functions. Especially one that binds all standardized otolith images together.
